New lower bounds for Schur and weak Schur numbers

### 2021-04-13 article/schur.tex
anglais

\qquad In this section, we use Rowley's constructions \cite{RowleyRamsey} in the context of Schur numbers. To improve lower bounds
for Ramsay's numbers, Rowley introduces partitions verifying some properties which can be extended using a method which generalizes 
Abbott and Hanson's  construction \cite{AbbottHanson}. Rowley named these partitions "templates", and we will keep this name in the 
entire article. We find suitable templates and use them to find new lower bounds for Schur numbers.

@@

\begin{proposition}
	Let \(n \in [\![2, +\infty]\!]\), we have :
	\[
	2S(n-1)+1 \leqslant S^+(n) \leqslant S(n)
	\]
\end{proposition}

@@

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template of \(k+1\) colors and lenght \(p\),
	and a partition of \(n\) sum-free subsets of \([\![1,q]\!]\) then there exists a partition of \(n+k\) sum-free subsets
	of \([\![1,pq+m_{k+1}-1]\!]\). \(m_{k+1}\) is the first number colored with the \(k+1\)-th color in the \(SF\)-template.
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geqslant S^+(k+1)S(n) + m_{k+1} - 1 \]
\end{corollary}

\begin{remark}
The following proposition can help improve the additive constant of a SF-template. Although it does not allow us to improve the SF-templates 
we have found, the analogous of this proposition for WSF-templates (see next section) allows to improve one of them.
\end{remark}

@@

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. 
They extend vertically a sum-free partition, and horizontally an other sum-free partition. This way each "block" 
acts like a security zone for the other one. Here, the horizontal partition is no longer to the side of the 
vertical one, but it occupies the column of the special color of the \(SF\)-template, i.e the one without the 
extra condition. We give the following example for \(p = 7, q = 4, n = 2 ~\textrm{and}~ k = 2\). \\

This shows the inequality \(S(2+2) \geqslant S^+(3)S(2) + 4\). The special color is blue and is first used
to color 5, hence \(4 = m_{3} - 1\). 
We prooved \(S^+(3) = 9\). The special color is blue.

@@

To compare, here is a classic Abbott and Hanson partition : \(S(2+2) \geqslant S(2)(2S(2)+1)+S(2) \). 
\begin{center}
\[
\begin{array}{|*{9}{c|}}
	\multicolumn{4}{c}{\overset{S(2)}{\overbrace{\rule{2.6cm}{0pt}}}} \\
	\hline
	\cellcolor{red} 1 & \cellcolor{green} 2 & \cellcolor{green} 3 & \cellcolor{red} 4 & \cellcolor{blue} 5 & \cellcolor{blue} 6 & \cellcolor{blue} 7 & \cellcolor{blue} 8 & \cellcolor{blue} 9\\
	\hline
	\cellcolor{red} 10 & \cellcolor{green} 11 & \cellcolor{green} 12 & \cellcolor{red} 13 & \cellcolor{yellow} 14 & \cellcolor{yellow} 15 & \cellcolor{yellow} 16 & \cellcolor{yellow} 17 & \cellcolor{yellow} 18\\
	\hline
	\cellcolor{red} 19 & \cellcolor{green} 20 & \cellcolor{green} 21 & \cellcolor{red} 22 & \cellcolor{yellow} 23 & \cellcolor{yellow} 24 & \cellcolor{yellow} 25 & \cellcolor{yellow} 26 & \cellcolor{yellow} 27\\
	\hline
	\cellcolor{red} 28 & \cellcolor{green} 29 & \cellcolor{green} 30 & \cellcolor{red} 31 & \cellcolor{blue} 32 & \cellcolor{blue} 33 & \cellcolor{blue} 34 & \cellcolor{blue} 35 & \cellcolor{blue} 36 \\
	\hline
	\cellcolor{red} 37 & \cellcolor{green} 38 & \cellcolor{green} 39 & \cellcolor{red} 40 \\
	\cline{1-4}
\end{array}
\]
\end{center}

>>>>

\qquad In this section, we use Rowley's constructions \cite{RowleyRamsey} in the context of Schur numbers. To improve lower bounds
for Ramsay's numbers, Rowley introduces some sum-free partitions verifying some additional properties which can be extended using a 
method which generalizes Abbott and Hanson's  construction \cite{AbbottHanson}. Rowley named these partitions "templates", and 
we will keep this name in the entire article. We find suitable templates and use them to find new lower bounds for Schur numbers.

@@

\begin{proposition}
	Let \(n \in [\![2, +\infty]\!]\). Then
	\[
	2S(n-1)+1 \leqslant S^+(n) \leqslant S(n)
	\]
\end{proposition}

@@

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template of \(k+1\) colors and lenght \(p\),
	and a partition of \(n\) sum-free subsets of \([\![1,q]\!]\) then there exists a partition of \(n+k\) sum-free subsets
	of \([\![1,pq+m_{k+1}-1]\!]\). \(m_{k+1}\) is the minimum of the special subset in the \(SF\)-template.
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S(n+k) \geqslant S^+(k+1)S(n) + m_{k+1} - 1 \]
\end{corollary}

\begin{remark}
The following proposition can help improve the additive constant of a SF-template. Although it does not allow us to improve the SF-templates 
we have found, the analogous of this proposition for WSF-templates (see next section) allows us to improve one of them.
\end{remark}

@@

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. 
They extend vertically a sum-free partition, and horizontally an other sum-free partition. This way each "block" 
acts like a security zone for the other one. Here, the horizontal partition is no longer to the side of the 
vertical one, but it occupies the column of the special color of the \(SF\)-template, i.e the one without the 
extra condition. We give the following example for \(p = 9, q = 4, n = 2 ~\textrm{and}~ k = 2\). \\

This shows the inequality \(S(2+2) \geqslant S^+(3)S(2) + 4\). The special color is blue and is first used
to color 5, hence \(4 = m_{3} - 1\). The special color is blue.

@@

To comparison, here is an usual Abbott and Hanson partition : \(S(2+2) \geqslant S(2)(2S(2)+1)+S(2) \). 
\begin{center}
\[
\begin{array}{|*{9}{c|}}
	\multicolumn{4}{c}{\overset{S(2)}{\overbrace{\rule{2.6cm}{0pt}}}} \\
	\hline
	\cellcolor{red} 1 & \cellcolor{green} 2 & \cellcolor{green} 3 & \cellcolor{red} 4 & \cellcolor{blue} 5 & \cellcolor{blue} 6 & \cellcolor{blue} 7 & \cellcolor{blue} 8 & \cellcolor{blue} 9\\
	\hline
	\cellcolor{red} 10 & \cellcolor{green} 11 & \cellcolor{green} 12 & \cellcolor{red} 13 & \cellcolor{yellow} 14 & \cellcolor{yellow} 15 & \cellcolor{yellow} 16 & \cellcolor{yellow} 17 & \cellcolor{yellow} 18\\
	\hline
	\cellcolor{red} 19 & \cellcolor{green} 20 & \cellcolor{green} 21 & \cellcolor{red} 22 & \cellcolor{yellow} 23 & \cellcolor{yellow} 24 & \cellcolor{yellow} 25 & \cellcolor{yellow} 26 & \cellcolor{yellow} 27\\
	\hline
	\cellcolor{red} 28 & \cellcolor{green} 29 & \cellcolor{green} 30 & \cellcolor{red} 31 & \cellcolor{blue} 32 & \cellcolor{blue} 33 & \cellcolor{blue} 34 & \cellcolor{blue} 35 & \cellcolor{blue} 36 \\
	\hline
	\cellcolor{red} 37 & \cellcolor{green} 38 & \cellcolor{green} 39 & \cellcolor{red} 40 \\
	\cline{1-4}
\end{array}
\]
\end{center}
